---
subtitle: "TMA4268 Statistical Learning V2022"
title: "Compulsory exercise 1: Group 5"
author: "Aleksander Johnsen Solberg and Gjermund Oscar Lyckander"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  # html_document
  pdf_document
---
  
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE,tidy=TRUE,message=FALSE,warning=FALSE,strip.white=TRUE,prompt=FALSE,
                      cache=TRUE, size="scriptsize",fig.width=4, fig.height=3,fig.align = "center")
```

```{r,eval=TRUE,echo=FALSE}
#install.packages("knitr") #probably already installed
#install.packages("rmarkdown") #probably already installed
#install.packages("ggplot2") #plotting with ggplot
#install.packages("palmerpenguins")
#install.packages("ggfortify") # For model checking
#install.packages("MASS")
#install.packages("class")
#install.packages("pROC")
#install.packages("plotROC")
#install.packages("boot")
library("knitr")
library("rmarkdown")
library("palmerpenguins")
```

<!--  Etc (load all packages needed). -->


# Problem 1

## a)
$$
\begin{aligned}
  E[y_0 - \hat f(x_0)]^2 &= E[(f(x_0) + \varepsilon - \hat f(x_0))^2] \\ 
  &= E[(f(x_0))^2] + E[\varepsilon^2] + E[\hat f(x_0)^2] -2E[f(x_0) \hat f(x_0)] + 2E[f(x_0) \varepsilon] + 2E[ \hat f(x_0) \varepsilon] \\
  &= f(x_0)^2 + \text{Var}(\varepsilon) + \text{Var}(\hat f(x_0)) + E[\hat f(x_0)]^2 - 2E[f(x_0) \hat f(x_0)]\\
  &= (f(x_0) - E[\hat f(x_0)])^2 + \text{Var}(\hat f(x_0))  + \text{Var}(\varepsilon)
\end{aligned}
$$
The three terms in the last line are the squared bias, variance, and irreducible error respectively.

## b)
The three terms can be interpreted as the following.
The bias term is the error that comes from modeling a complicated real-life problem with a simple model. The more flexible the model is, the smaller the bias will be. 
The variance term is how much the estimate $\hat{f}$ would change if we were using different training data. The more flexible the model is, the larger the variance will be.
Lastly, the irreducible error term is simply the error that comes from the error in the data itself.

## c)

## d)

## e)

 

# Problem 2

Here is a code chunk:

```{r, eval=TRUE}
library(palmerpenguins) # Contains the data set "penguins".
data(penguins)
head(penguins)
```


## a)

## b) 
 
## c)


# Problem 3

# Problem 4
 